---
title: "What's cooking"
subtitle: "Predictive modelling excercise"
author: "Silke Johannsen"
date: '2019-03-01'
output:
  html_document:
    keep_md: yes
    toc: yes
---


```{r global_Knitr_options, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

## Global knitr setting used in every chunk

knitr::opts_chunk$set(cache= FALSE, 
                      # Turn cache=TRUE if knitting the document is too slow
                      results='asis',
                      warning=FALSE,
                      message=FALSE,
                      echo=FALSE,
                      dpi=96,
                      fig.width=10,
                      fig.height=10, 
                      # Default figure widths
                      dev="png",
                      # dev.args=list(type="cairo"), # The png device
                      # Change to dev="postscript" if you want the EPS-files
                      # for submitting. Also remove the dev.args() 
                      # as the postscript
                      # doesn't accept the type="cairo" argument.
                      error=FALSE)
 
# Evaluate the figure caption after the plot
knitr::opts_knit$set(eval.after='fig.cap')

## NA's are rendered blank when printing a table
options(knitr.kable.NA = '')

## printed numbers rounded with big marks
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})

```

```{r versions}

info <- sessionInfo()
r_ver <- paste(info$R.version$major, info$R.version$minor, sep=".")
r_year <- info$R.version$year

```

## R information  
All analyses were performed using R (ver. `r r_ver`) [R Core Team, `r r_year`].

```{r set_up}

library(tidyverse)
library(here)
library(jsonlite)
library(kableExtra)
library(skimr)

```

## Load & inspect the data

The data set from the Kaggle competition, What's cooking is stored in JSON format. Load the train and test data with the jsonlite package and inspect the objects.

<br>
```{r read_data}

## read data

train_raw_df <- 
  read_json(here("data", "train.json"),
            simplifyVector = TRUE)

test_raw_df <- 
  read_json(here("data", "test.json"),
            simplifyVector = TRUE)

```


```{r inspec_data}

## inspecting the objects

head(train_raw_df) %>% 
  knitr::kable(caption = "Train data frame:") %>% 
  kable_styling()

head(test_raw_df) %>% 
  knitr::kable(caption = "Test data frame:") %>% 
  kable_styling()

## ckecking id and cuisine

train_no_id <- train_raw_df %>% 
  distinct(id) %>% 
  tally() %>% 
  pull()

train_no_cuisine <- train_raw_df %>% 
  distinct(cuisine) %>% 
  tally() %>% 
  pull()

test_no_id <- test_raw_df %>%
  distinct(id) %>% 
  tally() %>% 
  pull()


## are id's in test data set found in the train data set

test_id_vector <- test_raw_df %>% 
  distinct(id) %>% 
  pull()

test_id_in_train <- train_raw_df %>% 
  filter(id %in% test_id_vector)

rm(test_id_vector)

```
<br>

The train data have three columns named id, cuisine and ingredients, with the ingredients column being a list. The variable id has no duplicates and `r train_no_id` unique entries. There are `r train_no_cuisine` distinct cuisine entries.

The test data have two columns, id and ingredients. Ingredients again as a list. The variable id has no duplicates and `r test_no_id` unique entries. None of the id's in the test data set are present in the train data set.

Using unnest from the tidyr package to extract the ingredients list into a more user friendly format.

<br>
```{r flatten_df}

train_flat <- train_raw_df %>%
  unnest(ingredients) 

head(train_flat) %>% 
  knitr::kable(caption = "Train data frame flat:", align = "ccc") %>% 
  kable_styling()

test_flat <- test_raw_df %>%
  unnest(ingredients)

head(test_flat) %>% 
  knitr::kable(caption = "Test data frame flat:", align = "ccc") %>% 
  kable_styling()

```
<br>

It is now easy to do some simple summary analysis on the data using the dplyr & skimr package.

<br>
```{r}

train_flat %>% 
  mutate_all(funs(as.factor(.))) %>% 
  skim_to_wide() %>%
  knitr::kable(caption = " Train data: Summary statistics") %>% 
  kable_styling()

train_flat %>% 
  group_by(ingredients) %>% 
  tally() %>% 
  top_n(10, n) %>% 
  arrange(desc(n)) %>% 
  knitr::kable(caption = "Train data: Top 10 Ingredients") %>% 
  kable_styling()


test_flat %>% 
  group_by(ingredients) %>% 
  tally() %>% 
  top_n(10, n) %>% 
  arrange(desc(n)) %>% 
  knitr::kable(caption = "Test data: Top 10 Ingredients") %>% 
  kable_styling()

```

